---
title: "Data Transfer and Anonymisation"
author: "John Marc"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
# Load required packages.
library(httr)
```

## Overview

This markdown file fetches the Citizen Science Beetle Records For Ireland from its host server, anonymises observer names, and uploads the anonymised dataset to IPFS via Pinata.

## Step 1: Fetch data from host

```{r echo=FALSE}
rawDataUrl <- "https://maps.biodiversityireland.ie/Dataset/Download?datasetId=393"
```

The raw data is hosted at `r rawDataUrl`.

1. Create a temporary file to store the download.

```{r}
destFile <- tempfile(fileext = ".zip")
```

2. Download the data into the local, temporary file.

```{r}
# Download the ZIP file
download.file(
  url = rawDataUrl, 
  destfile = destFile, 
  mode = "wb")
```

3. Extract the data and metadata from the downloaded ZIP archive.

```{r}
# Directory where files will be extracted to
dirExtractTo <- "CitizenScienceBeetleRecords"
dir.create(dirExtractTo, showWarnings = FALSE)

# Unzip the archive
unzip(zipfile = destFile, exdir = dirExtractTo)
```

The ZIP archive contained the following files:

```{r}
# List extracted files
list.files(dirExtractTo, recursive = TRUE)
```

The CitizenScienceBeetleRecords.txt file contains the actual data; the other files are metadata files.

4. Load the data into R.

We'll use read.csv() with the the tab character as the separator.

```{r}
fnmData <- "CitizenScienceBeetleRecords.txt"
fnmsMetadata <- setdiff(list.files(dirExtractTo, recursive = TRUE), fnmData)
dfData <- read.csv(file = paste0(dirExtractTo, "/", fnmData), sep = "\t")
```

## Step 2: Encrypt sensitive variables

For the purposes of demonstration, we shall pretend that the names of the observers, as recorded in the 'Recorder' variable, should not be made public and should therefore be encrypted. We'll encrypt them by replacing them with their MD5 hashes.

```{r}
nUniqueBefore <- length(unique(dfData$Recorder))
dfData$Recorder <- sapply(X = dfData$Recorder, FUN = digest::digest, algo = "md5")
nUniqueAfter <- length(unique(dfData$Recorder))

if (nUniqueAfter != nUniqueBefore) {
  stop("Nonces required to preserve uniqueness in 'Recorder' during encryption.")
}
```

## Step 3: Upload the anonymised data and unmodified metadata to a content-addressed storage system

Note: Good practice is to upload the data and metadata as a CAR, so that the metadata always travels with the data. For the purposes of this demonstration, however, we are making use of Pinata's free plan, which does not support CAR creation. We shall therefore upload the data and the metadata as separate files.

1. Write the anonymised data to a CSV file.

```{r}
fpDataAnon <- tempfile(fileext = ".csv")
write.csv(x = dfData, file = fpDataAnon)
```

2. Upload the CSV containing the anonymised data to IPFS, via Pinata.

```{r}
res <- httr::POST(
  url = "https://uploads.pinata.cloud/v3/files", 
  config = httr::add_headers(
    .headers = c("Authorization" = paste("Bearer", Sys.getenv("PINATA_JWT")))), 
  body = list(
    network = "public",
    file = httr::upload_file(path = fpDataAnon)), 
  encode = "multipart")

if (res$status_code >= 300) {
  stop("res$status_code = ", res$status_code)
}

res <- httr::content(x = res, as = "parsed")
cidData <- res$data$cid
```

IPFS CID of anonymised data file: `r cidData`.

3. Upload the metadata files to IPFS, via Pinata.

```{r}
metadataCids <- sapply(X = fnmsMetadata, FUN = function(fnm) {
  
  fp <- paste0(dirExtractTo, "/", fnm)
  
  res <- httr::POST(
    url = "https://uploads.pinata.cloud/v3/files", 
    config = httr::add_headers(
      .headers = c("Authorization" = paste("Bearer", Sys.getenv("PINATA_JWT")))), 
    body = list(
      network = "public",
      file = httr::upload_file(path = fp)), 
    encode = "multipart")
  
  if (res$status_code >= 300) {
    stop("res$status_code = ", res$status_code)
  }
  
  res <- httr::content(x = res, as = "parsed")
  cid <- res$data$cid
  cat(sprintf("IPFS CID for metadata file %s: %s\n", fnm, cid))
  
  return(cid)
  
})
```

## R session info

```{r child="session-info.Rmd"}
```

