---
title: "Data Cleaning"
author: "John Marc"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This markdown file fetches the anonymised dataset (produced by [1-transfer-and-anonymisation.Rmd](#TODO-URLhere)) and takes it through the second phase of the Jellyfish data processing approach. The second phase involves cleaning (quality control) and transformation (if necessary).

In the interest of keeping the demonstration simple and easy to follow, the quality control performed by this file will be kept relatively minimal.

## Step 1: Fetch anonymised data from content-addressed storage

```{r}
anonDataCid <- "bafybeigvdwaevrt2kugd54xlqra6gwaiqygthzbaq26fm4e573jyoxjqzi"
ipfsGatewayUrl <- "https://ipfs.io/ipfs"
```

The IPFS CID of the anonymised data in CSV format is `r anonDataCid`.

```{r}
fpDownload <- tempfile(fileext = ".csv")
download.file(
  url = paste0(ipfsGatewayUrl, "/", anonDataCid), 
  destfile = fpDownload, 
  mode = "wb")
```

## Step 2: Read the data into a data frame

```{r}
dfData <- read.csv(fpDownload)
```

## Step 3: Standardise variable names

```{r}
names(dfData)[names(dfData) == "Vice.county"] <- "ViceCounty"
names(dfData)[names(dfData) == "Additional.information"] <- "AdditionalInformation"
names(dfData) <- snakecase::to_snake_case(names(dfData))
```

## Step 4: Cleaning and transformation: First pass

### Replace all empty character values with NA_character_

Note: Variable 'site_name' seems to contain some multibyte characters and will therefore need to be dealth with separately.

```{r}
idxxChar <- which(sapply(X = dfData, FUN = is.character))

for (vnm in setdiff(names(dfData)[idxxChar], "site_name")) {
  x <- dfData[[vnm]]
  x[which(nchar(x) == 0)] <- NA_character_
  dfData[[vnm]] <- x
}
```

### Variable 'x'

This variable was added automatically when we wrote the anonymised data to CSV in 1-transfer-and-anonymisation.Rmd, so we should remove it.

```{r}
dfData$x <- NULL
```

### Variable 'record_key'

This variable gives the ID of each record and should therefore be a character field to prevent it from being modified by accident during numeric operations.

```{r}
dfData$record_key <- format(dfData$record_key, scientific = FALSE)
```

### Variable 'survey_key'

Check what's inside:

```{r}
table(dfData$survey_key, exclude = NULL)
```

This is an empty variable and can therefore be removed:

```{r}
dfData$survey_key <- NULL
```

### Variable 'sample_key'

Check what's inside:

```{r}
table(dfData$sample_key, exclude = NULL)
```

This is an empty variable and can therefore be removed:

```{r}
dfData$sample_key <- NULL
```

### Variables 'start_date', 'end_date', 'date_type', and 'date'

Inspection:

```{r}
table(dfData$date_type, exclude = NULL)
```

Variable 'date_type' is not informative and can therefore be removed:

```{r}
dfData$date_type <- NULL
```

Convert 'start_date', 'end_date' and 'date' to type 'Date', being careful not to lose any values due to textual entry errors:

```{r}
for (vnm in c("start_date", "end_date", "date")) {
  nNaBefore <- length(which(is.na(dfData[[vnm]])))
  dfData[[vnm]] <- as.Date.character(x = dfData[[vnm]], format = "%d/%m/%Y")
  nNaAfter <- length(which(is.na(dfData[[vnm]])))
  if (nNaAfter > nNaBefore) {
    stop(sprintf("NAs introduced for '%s' through character-to-date conversion.", vnm))
  }
}
```

According to the dataset's metadata, all dates should be between 2006 and 2024 (both years included), so we'll check for that:

```{r}
dfDateRanges <- as.data.frame(
  data.table::rbindlist(
    lapply(X = c("start_date", "end_date", "date"), FUN = function(vnm) {
      data.frame(
        var = vnm, 
        min_yr = min(lubridate::year(dfData[[vnm]])),
        max_yr = max(lubridate::year(dfData[[vnm]])))
    })))

knitr::kable(dfDateRanges)
```

All dates seem within range.

Check that 'date' is either the same as 'start_date' or the same as 'end_date':

```{r}
idxx <- which(dfData$date != dfData$start_date & dfData$date != dfData$end_date)
```

No records (`r length(idxx)`) have a date value different from both their start_date and their end_date values, so all seems well.

### Variable 'taxon_version_key'

Check values:

```{r}
table(dfData$taxon_version_key, exclude = NULL)
```

Nothing suspicious.

### Variable 'taxon_name'

Check values:

```{r}
table(dfData$taxon_name, exclude = NULL)
```

Nothing suspicious.

### Variable 'grid_reference'

This gives the location of the observation according to the Irish National Grid projection. Let's check if all values follow the same spatial resolution:

```{r}
table(nchar(dfData$grid_reference))
```

Different spatial resolutions are represented, so we need to convert all references to the same spatial resolution. The lowest, common resolution seem to be 1 km (5 characters, i.e., 1 letter + 2 digits).

```{r}
idxx <- which(nchar(dfData$grid_reference) > 5)
dfData$grid_reference[idxx] <- sapply(X = dfData$grid_reference[idxx], FUN = function(x) {
  splts <- strsplit(x = x, split = "")[[1]]
  idxHalf <- 1 + (length(splts) -1)/2
  newX <- paste(c(splts[1], splts[2:3], splts[(idxHalf+1):(idxHalf+2)]), 
                sep = "", collapse = "")
  return(newX)
})
```

Check character lengths again:

```{r}
table(nchar(dfData$grid_reference))
```

All good.

### Variables 'east', 'north', and 'projection'

Some coordinates are given as OSI, while others are given as WGS84. Convert all to WGS84.

```{r}
idxx <- which(dfData$projection == "OSI")
pts_osi <- sf::st_as_sf(
  x = dfData[idxx,], 
  coords = c("east", "north"), 
  crs = 29902)  # 29902 or 2157
pts_wgs84 <- sf::st_transform(
  x = pts_osi, 
  crs = 4326)
#sf::st_coordinates(pts_wgs84)
#plot(pts_wgs84$geometry)

dfData$east[idxx] <- sf::st_coordinates(pts_wgs84)[,"X"]
dfData$north[idxx] <- sf::st_coordinates(pts_wgs84)[,"Y"]
dfData$projection[idxx] <- "WGS84"

plot(
  sf::st_as_sf(
    x = dfData, 
    coords = c("east", "north"), 
    crs = 4326)$geometry)
```

### Variable 'zero_abundance'

Check values:

```{r}
table(dfData$zero_abundance, exclude = NULL)
```

Convert to type 'logical':

```{r}
dfData$zero_abundance <- dfData$zero_abundance != "False"
```

### Variable 'under_validation'

Check values:

```{r}
table(dfData$under_validation, exclude = NULL)
```

Convert to type 'logical':

```{r}
dfData$under_validation <- dfData$under_validation != "False"
```

### Variable 'county':

Check values:

```{r}
table(dfData$county, exclude = NULL)
```

Nothing suspicious.

## Step 5: Upload cleaned data to IPFS

1. Save the data as an RDS file to preserve R-specific types and formatting.

```{r}
fpRds <- "dfCleaned.rds"
saveRDS(object = dfData, file = fpRds)
```

2. Upload the RDS to IPFS, via Pinata.

```{r}
res <- httr::POST(
  url = "https://uploads.pinata.cloud/v3/files", 
  config = httr::add_headers(
    .headers = c("Authorization" = paste("Bearer", Sys.getenv("PINATA_JWT")))), 
  body = list(
    network = "public",
    file = httr::upload_file(path = fpRds)), 
  encode = "multipart")

if (res$status_code >= 300) {
  stop("res$status_code = ", res$status_code)
}

res <- httr::content(x = res, as = "parsed")
cidData <- res$data$cid
```

IPFS CID of clean data RDS file: `r cidData`.

## R session info

```{r child="session-info.Rmd"}
```

