---
title: "Aggregation and Calculation"
author: "John Marc"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This markdown file fetches the cleaned dataset (produced by [2-cleaning.Rmd](#TODO-URLhere)) and calculates the variables required for reporting. The variables that will be calculated are those prescribed by the [Methodology for the Estimation of Beetle Biodiversity from Citizen Science Data](https://doesnotexist.xyz).

## Step 1: Fetch cleaned data from content-addressed storage

```{r}
cleanedDataCid <- "bafkreicakzjr66xkin4qbeu7sox32dysweuu2d5vdw4cfpllqmozfphnh4"
ipfsGatewayUrl <- "https://ipfs.io/ipfs"
```

The IPFS CID of the cleaned data in RDS format is `r cleanedDataCid`.

```{r}
fpDownload <- tempfile(fileext = ".rds")
download.file(
  url = paste0(ipfsGatewayUrl, "/", cleanedDataCid), 
  destfile = fpDownload, 
  mode = "wb")
```

## Step 2: Read the data into a data frame

```{r}
dfData <- readRDS(fpDownload)
```

## Step 3: Drop rows for years that do not meet the methodology's coverage requirements

The methodology requires at least 20 observations per year, when performing annual calculations.

```{r}
dfData$year <- lubridate::year(dfData$date)
```

The dataset currently contains the following number of observations for each year:

```{r}
dfNobsYr <- as.data.frame(
  dplyr::summarise(
    dplyr::group_by(dfData, year),
    n = dplyr::n(),
    .groups = "drop"))

knitr::kable(dfNobsYr)
```

Drop observations from years with fewer than 20 observations in total:

```{r}
dfData <- dfData[which(dfData$year %in% dfNobsYr$year[which(dfNobsYr$n >= 20)]),]
```

## Step 4: Descriptive statistics

Number of observations per taxon per year:

```{r}
dfNobsTaxYr <- as.data.frame(
  dplyr::summarise(
    dplyr::group_by(dfData, taxon_name, year),
    n = dplyr::n(),
    .groups = "drop"))

knitr::kable(x = dfNobsTaxYr[1:10,])
```

Average number of observations per taxon per year:

```{r}
dfAvgNobsTaxYr <- as.data.frame(
  dplyr::summarise(
    dplyr::group_by(dfNobsTaxYr, taxon_name),
    n_avg = round(mean(n), digits = 2),
    .groups = "drop"))

knitr::kable(x = dfAvgNobsTaxYr[1:10,])
```

Number of taxa observed per county per year:

```{r}
dfNtaxCountyYr <- as.data.frame(
  dplyr::summarise(
    dplyr::group_by(dfData, county, year),
    n = length(unique(taxon_name)),
    .groups = "drop"))

knitr::kable(x = dfNtaxCountyYr[1:10,])
```

## Step 5: Calculate variables required by methodology

Average number of taxa observed per square kilometre per year:

```{r}
dfNtaxSqKmYr <- as.data.frame(
  dplyr::summarise(
    dplyr::group_by(dfData, year),
    n_taxa = length(unique(taxon_name)),
    n_sq_km = length(unique(grid_reference)),
    .groups = "drop"))

dfNtaxSqKmYr$ntax_p_sqkm <- dfNtaxSqKmYr$n_taxa / dfNtaxSqKmYr$n_sq_km

knitr::kable(dfNtaxSqKmYr)
```

## Step 6: Store calculated values on IPFS

1. Jellyfish strongly recommends storing the calculation results as JSON, so let's write the results to a JSON file:

```{r}
fpOut <- "calculationResults.json"
jsonOut <- jsonlite::write_json(dfNtaxSqKmYr, path = fpOut)
```

2. Upload the JSON file to IPFS, via Pinata.

```{r}
res <- httr::POST(
  url = "https://uploads.pinata.cloud/v3/files", 
  config = httr::add_headers(
    .headers = c("Authorization" = paste("Bearer", Sys.getenv("PINATA_JWT")))), 
  body = list(
    network = "public",
    file = httr::upload_file(path = fpOut)), 
  encode = "multipart")

if (res$status_code >= 300) {
  stop("res$status_code = ", res$status_code)
}

res <- httr::content(x = res, as = "parsed")
cidCalcRes <- res$data$cid
```

IPFS CID of the calculation results as JSON: `r cidCalcRes`.

## R session info

```{r child="session-info.Rmd"}
```

